AI-Based Detection of Jamming Attacks In 6G Drone Networks

### 2024-09-26 article.tex
Wrote related works

%PLACE TO ADD COMMANDS

@@

% Authors, for the paper (add full first names)
\Author{Sergio Cibecchini $^{1,\dagger,\ddagger}$\orcidA{}, Francesco Chiti $^{2,\ddagger}$ and Firstname Lastname $^{2,}$*}

@@

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Sergio Cibecchini, Francesco Chiti and Firstname Lastname}

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusions: indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

@@

The shift from 4G to 5G has been a generational leap has revolutionized connectivity around the world. 
Despite 5G being still in its early adoption rate at the time of writing \cite{5GStatisticsTaylor}, the research community is already working on the next generation of wireless communication technologies, 6G. 
6G technology is expected to enable a wide variety of new use cases, thanks to the increases in both data rates and latency but this in turn 
will also bring forth new security challenges, specific to those applications.

One of the main differences between 5G and 6G technology will be the focus of the 6G standard in regards to 
AI integration. While Software Defined Networks (SDN) have played a key role in improving the efficiency and security 
of 5G networks, 6G is expected to take this a step further, by integrating artificial
intelligence and machine learning directly into the network. This is what the authors of \cite{6GRoadmapLetaief} define as 
the shift from \textit{Softwarization} to \textit{Intelligentization}.

AI integration in 6G networks will greatly strengthen the security of the network against security threats. 
By leveraging Diagnostic Analytics, a collection of insights into the status of the networks, security teams 
will be able to train specific AI models to detect and respond to security threats in real-time.

@@

The paper is divided in 5 sections: 
\begin{enumerate}
	\item \textbf{Topic overview:} In this section we will discuss how drones can benefit from integration in 6G network, analyze jamming attacks, their types, mitigation and detection. Finally we will discuss the advantages of an edge AI approach for jamming detection. 
	\item \textbf{Materials and methods:} In this section we will define the scenario that we decided to analyze as well as the dataset, algorithm and evaluation metrics we chose for our tests. 
	\item \textbf{Numerical Results:} In this section we will present the testing methodology as well as the numerical results that we obtained. 
	\item \textbf{Discussion:} In this section we will discuss the results obtained in the previous section and highlight notable trends.
	\item \textbf{Conclusions:} In this section we will summarize the results and discuss possible future research directions.
\end{enumerate}

@@

%TODO Add related works

@@

In report 22.886 \cite{5GV2XSultan} 3GPP, the organ responsible for the development and maintenance of the 5G standard, identifies some of the envisioned use cases for 5G V2X (Vehicle-to-
Everything) communication services. Among these use cases, the report identifies vehicle platooning, advanced and remote driving end extended situational awareness as some of the main benefits
of V2V communication. All these use cases can be leveraged by a 6G drone network to achieve fast and reliable drone-to-drone communication, that, with the integration of artificial intelligence 
would allow the drones to act autonomously in a coordinated manner.

This would prove useful in a variety of fields: from autonomous soil and crop health assessment as well as irrigation in agriculture, 
delivery of life saving supplies and locating of survivors in disaster response, in the delivery sector as a more eco friendly alternative to traditional 
delivery options and possibly in the mobility sector as a complement to traditional taxis and public transportation \cite{DroneCommHassija}.

The effectiveness of drones in the military sector is widely recognized, both for high and low end models. For example, in the Ukraine war even cheap FPV drones 
mounted with explosives were successfully used to destroy much more expensive equipment \cite{DroneCombatUkraine}.

@@

\begin{itemize}
    \item \textbf{Constant Jamming:} Constant jamming is the simplest form of jamming attack. In constant jamming, the jammer is active all the time. 
    \item \textbf{Periodic Jamming:} The jammer simply transmits a strong signal on the same frequency as the devices it wants to disrupt 
    for a certain period of time \(t_a\), then stops transmitting for another period of time \(t_b\). This type of jamming attack is particularly effective against devices 
    that are not able to change their frequency and has the ability to disrupt the transmission of a sequence of consecutive packets. \cite{VANETsAI-Lyamin}
    \item \textbf{Random Jamming:} In random jamming, the jammer is active at random intervals, giving each transmitted packet a probability 
    \(p\) of being jammed \cite{VANETsAI-Lyamin}. 
    \item \textbf{Reactive Jamming:} A reactive jammer starts transmitting its jamming signal only when it senses energy in the communication channel, 
    indicating that a legitimate transmission is taking place. This type of jamming attack is more power efficient compared to other 
    operation types, as it only transmits its signal when it is needed\cite{MLMisbehavior5GBoualouache}.
    \item \textbf{Smart Jamming:} A smart jammer is a more sophisticated type of jammer that is able to adapt its jamming signal to 
    maximize the disruption of the communication between the devices. Smart jammers are able to modify their attack pattern based on the 
    transmission specifics of the devices they are targeting and are able to adapt to changes in the communication channel\cite{AntiJammingV2V-Feng}.
\end{itemize}

@@

\subsection{Jamming attacks against drone networks}

@@

One real world example of this is the capture of an American drone by Iran in 2011. 
In December 2011 a Lockheed Martin RQ-170 Sentinel drone operated by the United States Air Force was flying over Iran when its operators lost control of the vehicle. The US government initially claimed that the drone had crashed due to a technical malfunction, but later reports revealed that the drone had been captured by the Iranian military. 
Iranian electronic warfare specialists claimed to have brought it down using a Jamming attack, that forced the drone into a return to base procedure, in combination with a  GPS spoofing attack, that made the drone land into a designated area \cite{RQ170DroneOwano}. 
After successfully capturing the drone, the Iranian government managed to reverse-engineer the drone and produce a working replica, which was then used in their military operations \cite{IranianUAVGross}.

@@

@@

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
\begin{quote}
This is an example of a quote.
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
\subsection{Subsection}
\subsubsection{Subsubsection}

The text continues here.

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, etc.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
\caption{This is a figure. Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.\label{fig1}}
\end{figure}   
\unskip

\begin{table}[H] 
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCC}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
Entry 1		& Data			& Data\\
Entry 2		& Data			& Data \textsuperscript{1}\\
\bottomrule
\end{tabularx}
\noindent{\footnotesize{\textsuperscript{1} Tables may have a footer.}}
\end{table}

The text continues here (Figure~\ref{fig2} and Table~\ref{tab2}).

% Example of a figure that spans the whole page width. The same concept works for tables, too.
\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=15.5cm]{Definitions/logo-mdpi}
\end{adjustwidth}
\caption{This is a wide figure.\label{fig2}}
\end{figure}

\begin{table}[H]
\caption{This is a wide table.\label{tab2}}
	\begin{adjustwidth}{-\extralength}{0cm}
		\newcolumntype{C}{>{\centering\arraybackslash}X}
		\begin{tabularx}{\fulllength}{CCCC}
			\toprule
			\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}     & \textbf{Title 4}\\
			\midrule
\multirow[m]{3}{*}{Entry 1 *}	& Data			& Data			& Data\\
			  	                   & Data			& Data			& Data\\
			             	      & Data			& Data			& Data\\
                   \midrule
\multirow[m]{3}{*}{Entry 2}    & Data			& Data			& Data\\
			  	                  & Data			& Data			& Data\\
			             	     & Data			& Data			& Data\\
                   \midrule
\multirow[m]{3}{*}{Entry 3}    & Data			& Data			& Data\\
			  	                 & Data			& Data			& Data\\
			             	    & Data			& Data			& Data\\
                  \midrule
\multirow[m]{3}{*}{Entry 4}   & Data			& Data			& Data\\
			  	                 & Data			& Data			& Data\\
			             	    & Data			& Data			& Data\\
			\bottomrule
		\end{tabularx}
	\end{adjustwidth}
	\noindent{\footnotesize{* Tables may have a footer.}}
\end{table}

%\begin{listing}[H]
%\caption{Title of the listing}
%\rule{\columnwidth}{1pt}
%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
%\rule{\columnwidth}{1pt}
%\end{listing}

Text.

Text.

\subsection{Formatting of Mathematical Components}

This is the example 1 of equation:
\begin{linenomath}
\begin{equation}
a = 1,
\end{equation}
\end{linenomath}
the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct line numbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.

This is the example 2 of equation:
\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
\end{equation}
\end{adjustwidth}

% Example of a page in landscape format (with table and table footnote).
%\startlandscape
%\begin{table}[H] %% Table in wide page
%\caption{This is a very wide table.\label{tab3}}
%	\begin{tabularx}{\textwidth}{CCCC}
%		\toprule
%		\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}	& \textbf{Title 4}\\
%		\midrule
%		Entry 1		& Data			& Data			& This cell has some longer content that runs over two lines.\\
%		Entry 2		& Data			& Data			& Data\textsuperscript{1}\\
%		\bottomrule
%	\end{tabularx}
%	\begin{adjustwidth}{+\extralength}{0cm}
%		\noindent\footnotesize{\textsuperscript{1} This is a table footnote.}
%	\end{adjustwidth}
%\end{table}
%\finishlandscape

Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

>>>>

%PLACE TO ADD COMMANDS
\usepackage{lipsum}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{subcaption}

@@

% Authors, for the paper (add full first names)
\Author{Sergio Cibecchini $^{1,\dagger}$\orcidA{}, Francesco Chiti $^{2}$ and Laura Pierucci$^{2,}$*}

@@

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Sergio Cibecchini, Francesco Chiti and Laura Pierucci}

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{}

@@

The shift from 4G to 5G has been a generational leap has revolutionized connectivity around the world. 
Despite 5G being still in its early adoption rate at the time of writing \cite{5GStatisticsTaylor}, the research community is already working on the next generation of wireless communication technologies, 6G.

6G technology is expected to enable a wide variety of new use cases, thanks to the increases in both data rates and latency but this in turn 
will also bring forth new security challenges, specific to those applications.

One of the main differences between 5G and 6G technology will be a much deeper integration with AI. While Software Defined Networks (SDN) have played a key role in improving the efficiency and security 
of 5G networks, 6G is expected to take this a step further, by designing the networks to be integrated with artificial intelligence from the ground up. This is what the authors of \cite{6GRoadmapLetaief} define as 
the shift from \textit{Softwarization} to \textit{Intelligentization}.

AI integration in 6G networks will greatly strengthen the security of the network against potential threats. 
By leveraging Diagnostic Analytics, a collection of insights into the status of the networks, security teams 
will be able to train specific AI models to detect and respond to security threats in real-time.

@@

The paper is divided in 5 sections: 
\begin{enumerate}
	\item \textbf{Topic overview:} In this section we will make a general overview of how drones can benefit from integration in 6G network, analyze jamming attacks, their types, mitigation and detection. Finally we will discuss the advantages of an edge AI approach for jamming detection. 
	\item \textbf{Materials and methods:} In this section we will define the scenario that we decided to analyze as well as the dataset, algorithm and evaluation metrics we chose for our tests. 
	\item \textbf{Numerical Results:} In this section we will present the testing methodology, describe model tuning and testing phases and present the results obtained. 
	\item \textbf{Discussion:} In this section we will comment the results obtained in the previous section and highlight the advantages of the proposed approach. 
	\item \textbf{Conclusions:} In this section we will show how the proposed solution fits into a real world scenario and discuss possible future research directions.
\end{enumerate}

@@

Detection of jamming attacks in wireless networks has been a topic of interest for the research community for many years. A wide variety of approaches have been proposed for the detection of jamming attacks 
in different types of networks, with most of the work focusing on the detection of jamming attacks in Vehicular Ad-Hoc Networks (VANETs)\cite{VANETsAI-Lyamin}.

This work build upon the research done in \cite{JammingDetectionIoT-Hussain}, where the authors propose a machine learning based approach for the detection of jamming attacks in IoT networks. 
The authors propose the use of Convolutional Neural Networks (CNNs) to detect jamming attacks in a IoT drone scenario. The authors show that the CNN model is able to achieve a high detection rate against constant jamming attacks, but it 
struggles in the classification of periodic jamming and normal traffic.

The authors of \cite{GPSSpoofingDetection-Zuo} propose an Isolation Forest based approach for the detection of GPS spoofing attacks, while the authors of \cite{HybridJammingDetection-Hong} successfully applied a combination of decision trees and Isolation forest to classify 
and detect jamming attacks in a mobile scenario.

We apply a similar solution to the one proposed in \cite{HybridJammingDetection-Hong} to the scenario of a 6G drone network, testing the model against the dataset proposed in \cite{JammingDetectionIoT-Hussain}.
Our approach reached a high detection rate against both constant and periodic jamming attacks, while also being able to classify normal traffic with a high degree of accuracy. We also focused on the tuning of the model's hyperparameters to fit our specific scenario and compared it 
against the default model. In the tuning phase, the resource constrained nature of the scenario was taken into a count for the selection of the more appropriate hyperparameters.

@@

In report 22.886 \cite{5GV2XSultan} 3GPP, the organ responsible for the development and maintenance of the 5G standard, identifies some of the envisioned use cases for 5G V2X (Vehicle-to-
Everything) communication services. Among these use cases, the report identifies vehicle platooning, advanced and remote driving end extended situational awareness as some of the main benefits
of V2V communication. All these capabilities can be leveraged by a 6G drone network to achieve fast and reliable drone-to-drone communication, that, with the integration of artificial intelligence 
would allow the drones to act autonomously in a coordinated manner.

This would prove useful in a variety of fields: from autonomous irrigation as well as soil and crop health assessment in agriculture, 
delivery of life saving supplies and identification of survivors in disaster response, in the delivery sector as a more eco friendly alternative to traditional 
delivery options and possibly in the mobility sector as a complement to traditional taxis and public transportation \cite{DroneCommHassija}.

In the military sector, the effectiveness of drones is widely recognized, both for high and low end models. For example, in the Ukraine war even cheap FPV drones 
mounted with explosives were successfully used to destroy much more expensive equipment \cite{DroneCombatUkraine}.

@@

\begin{itemize}
    \item \textbf{Constant Jamming:} The jammer continuously transmits a strong signal on the same frequency used by the devices it wants to disrupt to communicate. 
    \item \textbf{Periodic Jamming:} The jammer transmits a strong signal for a certain period of time \(t_a\), then stops transmitting for another period of time \(t_b\). This pattern is repeated until the attack is stopped.
    \item \textbf{Random Jamming:} In random jamming, the jammer is active at random intervals, jamming each transmitted packet with a probability 
    \(p\) based on the attack pattern\cite{VANETsAI-Lyamin}. 
    \item \textbf{Reactive Jamming:} A reactive jammer starts transmitting its jamming signal only when it senses energy in the communication channel, 
    indicating that a legitimate transmission is taking place. This type of jamming attack is more power efficient compared to other 
    attack patterns, as it only transmits its signal when it know that it can actually disrupt the communication\cite{MLMisbehavior5GBoualouache}.
    \item \textbf{Smart Jamming:} A smart jammer is a more sophisticated type of jammer that is able to adapt its jamming signal to 
    maximize the disruption of the communication between the devices. Smart jammers are able to modify their attack pattern based on the 
    transmission specifics of the devices they are targeting and are able to adapt to changes in the communication channel\cite{AntiJammingV2V-Feng}.
\end{itemize}

@@

\subsection{Jamming attacks against drone networks}\label{JammingDroneNetworks}

@@

One real world example is the capture of an American drone by Iran in 2011. 
In December 2011 a Lockheed Martin \textit{RQ-170} Sentinel drone operated by the United States Air Force was flying over Iran when its operators lost control of the vehicle. The US government initially claimed that the drone had crashed due to a technical malfunction, but later reports revealed that the drone had been captured by the Iranian military. 
Iranian electronic warfare specialists claimed to have brought it down using a Jamming attack, that forced the drone into a return to base procedure, in combination with a  GPS spoofing attack, that made the drone land into a designated area \cite{RQ170DroneOwano}. 
After successfully capturing the drone, the Iranian government managed to reverse-engineer the drone and produce a working replica, which was then used in their military operations \cite{IranianUAVGross}.

@@

When presenting a machine learning based approach in an IoT setting, the question of where the AI model should be placed often arises. By their nature, IoT devices, and in turn drones, are usually resource constrained, 
both in terms of computational power but also in terms of internal storage and battery capacity \cite{6GSecurity-Chorti}. This means that complex AI models and algorithms are usually not feasible to be run on the device itself. A centralized approach 
offloads the computational burden to a central server, that returns the results of the AI model to the device. While this approach might be favorable in some cases, in a real-time situation such as a jamming attack, the latency introduced by the
communication with the central server means less time to react to the attack. Also, as jamming attacks degrade or sometimes completely block the communication between the device and the base station, a centralized approach might not be feasible in this case.
An lightweight edge approach on the other hand, while not as powerful as a centralized approach, would to provide real-time results to the device and would be able to operate even when the communication is disrupted, which is crucial in a jamming situation.

\subsection{Detection and Mitigation of Jamming Attacks}\label{sec:DetectionMitigation}

When attempting to mitigate jamming attacks, the first step is always to detect that an attack is taking place. This can be done by monitoring the communication channel for signs of jamming. 
The state of the art approaches involve the analysis of metrics such as the Signal to Noise Ration (SNR), the Received Signal Strength (RSS) and the Packet Delivery Ratio (PDR) \cite{JammingDetection-Sciancalepore}. The simplest approach when 
developing a method to detect jamming attacks is to set a static threshold for these metrics and trigger an alarm when the threshold is crossed. While this approach is easy to implement and might be effective in some cases, it is not able to adapt to changes in the communication channel and might be prone to false positives.
Implementing a machine learning based approach for jamming detection would allow the system to adapt to the changes in the communication channel and would be able to provide a more accurate detection of jamming attacks \cite{VANETsAI-Lyamin}. This is especially useful in mobile scenarios like 
drone networks, where the environment is constantly changing and the signal strength can vary greatly.

Once an attack is successfully detected, state of the art jamming mitigation techniques, like Direct Sequence Spread Spectrum (DSSS), Frequency Hopping (FHSS) and advanced signal processing techniques can be put in place to mitigate the 
effects of the attack.

The use of an AI model for jamming detection is particularly suitable for a 6G network, as the improved sensing capabilities would allow the model to train on a wide variety of diagnostic data, improving the detection rate and accuracy of the model. 
Furthermore, as the network is expected to be integrated with AI from the ground up, its highly likely that in the future drones will be integrated with AI-specific chips for a more efficient execution of AI models.

@@

\subsection{Scenario Definition and Attacker Classification}

In the proposed scenario we have a flying drone that is subject to a jamming attack. The drone communicates with a ground station and periodically samples the Received Signal Strength (RSS) of the signal it receives. 
The drone implements a simple unsupervised machine learning algorithm module, trained on nominal traffic RSS values. The module takes the sampled RSS values and classifies them as either normal or anomalous. Thanks to this classification module, the drone is able to determine wether or not a jamming attack is taking place.

We assume that a jammer is deployed in the network and that it starts a jamming attack at a certain point in time by transmitting a strong signal on the same frequency as the one used by the drone to communicate with the ground station.
We assume that the jammer is able to completely block the communication between the drone and the ground station, meaning the drone has to rely on its internal classification algorithm to determine if a jamming attack is taking place and react accordingly.
We assume that the jammer type is unknown to the drone.

Figure \ref{fig:CombinedJammingscenariosDiagram} shows a diagram of the proposed scenario.

\begin{figure}[H]
	\includegraphics[width=10.5 cm]{img/CombinedJammingscenariosDiagram.jpg}
	\caption{Proposed jamming scenario. The Drone communicates with the 6G base station and receives information about how to behave. The fixed jammer \textbf{(1)} and the mobile jammer \textbf{(2)} target the communication between the drone and the base station. The drone has an internal classification module that is able to detect jamming attacks.}
	\label{fig:CombinedJammingscenariosDiagram}
	\end{figure}   
	\unskip

In our scenario, the drone will be subject to two types of jamming attacks, a constant jamming attack and a periodic jamming attack.
These types of attacks can be attributed to different types of jammers. The constant jamming attack could be caused by a ground jammer, as maintaining a constant jamming signal requires a large power source, while 
the periodic jamming attack could be caused by a drone jammer, as drones are usually battery powered and can only jam for a limited amount of time. 
Employing a periodic jamming attack could help the malicious drone conserve energy and reduce the chances of being detected.

Table \ref{tab:attacker_classification} shows the attacker classification in the proposed scenario \cite{MLMisbehavior5GBoualouache}:

\begin{table}[H]
	\caption{Attacker classification details.\label{tab:attacker_classification}}
	\newcolumntype{C}{>{\centering\arraybackslash}X}
	\begin{tabularx}{\textwidth}{CC}
	\toprule
	\textbf{Classification} & \textbf{Description} \\
	\midrule
	Active   & The attacker is actively trying to disrupt network operation by transmitting a jamming signal\\
	External & The attack takes place at level 1 of the OSI model, meaning that the attacker is not part of the network\\
	Local    & The attack is local, as it is targeted at a specific drone or drone cluster and not at the entire network\\
	Malicious & Jamming attacks are considered malicious as their main goal is to disrupt correct network operation\\
	\bottomrule
\end{tabularx}
\end{table}

\subsection{Dataset Choice}\label{sec:DatasetChoice}

As 6G network traffic is not yet available, we chose an open-source dataset \cite{JammingDetectionIoT-Hussain} that analyses the RSS values 
received by a Raspberry Pi 3 that is subject to periodic and constant jamming attacks.

The dataset was created using a software-defined radio (SDR) connected to a laptop that was programmed to transmit a jamming signal using the 
open source software \textit{GNU Radio}. A second SDR radio was connected to a Raspberry Pi 3 and was chosen to receive the jamming signal. The jamming radio operates at a frequency of $2.412 GHz$ with a Bandwidth of $40MHz$, while the Raspberry Pi 3 was programmed to sample the RSS values with a frequency of $32K$ samples per second.

The dataset contains 3 different \textit{.txt} files that store the RSS values sampled by the Raspberry Pi 3. The first file contains the RSS values sampled during a constant jamming attack, the second file contains the RSS values sampled during a periodic jamming attack and the third file contains the RSS values sampled during normal network operation.
The samples are stored in a single column, with each row representing a single sample. The samples us the \textit{dBm (decibel-milliwatts)} unit of measurement, a logarithmic scale that measures the power of a signal in relation to 1 milliwatt. 
Conversion can be done using the formula \cite{WirelessCommSobot} shown in equation \ref{eq:dbm_to_mw}:

\begin{linenomath}
	\begin{equation}
		P_{dBm} = 10 \cdot \log_{10} \left( \frac{P_{mW}}{1mW} \right)
		\label{eq:dbm_to_mw}
	\end{equation}
\end{linenomath}

Figures \ref{fig:ConstantJammingSignal} and \ref{fig:PeriodicJammingSignal} show respectively plots of the RSS values sampled during a constant jamming attack and a periodic jamming attack.

\begin{figure}[H]
	\includegraphics[width=10.5 cm]{img/ConstantJammingSignal.png}
	\caption{Constant jamming attack RSS values}
	\label{fig:ConstantJammingSignal}
\end{figure}   
\unskip
\begin{figure}[H]
	\includegraphics[width=10.5 cm]{img/PeriodicJammingSignal.png}
	\caption{Periodic jamming attack RSS values}
	\label{fig:PeriodicJammingSignal}
\end{figure}   
\unskip

\subsection {Algorithm Choice}

When choosing the algorithm for the classification module, we first had to define the requirements that the algorithm had to meet. The algorithm has to be able to classify anomalous samples with a high detection rate (Recall), while at the same time 
being lightweight enough to be run on a resource constrained device like a drone. The training phase should also be fast, as this would allow the drone to redetermine the normal transmission RSS values in a constantly changing environment. If the model didn't periodically redefine normal RSS values, 
it would be prone to false positives caused by the mobile nature of drones. 
The model should be also required a small amount of storage to run effectively, as drones are usually very limited in internal memory.

The algorithm that best fit these requirements was the \textit{Isolation Forest} algorithm \cite{IsolationForestLiu}. The Isolation Forest algorithm is a state-of-the-art unsupervised machine learning algorithm for anomaly detection, introduced by Liu et al. in 2008. 
Unlike traditional anomaly detection algorithms, which require the definition of a normal class, Isolation Forest is able to detect anomalies without explicitly defining its characteristics.
This is achieved by leveraging the properties of anomalies themselves, i.e being few and separated from the majority of the data points.

The algorithm works by building a forest of isolation trees. Each tree is built by randomly selecting one of the features and then splitting the data points based on a value randomly selected between the minimum and maximum value of the feature. This partitions the data into the 
left and right branches of the tree. The process is repeated recursively until all the data points are isolated. The height of a node, meaning the number of splits that were required to isolate it, is used to determine the anomaly score of the data point. The higher the number of splits, the more likely the data point is to be an anomaly.
Isolation Forest is an \textit{ensamble} algorithm, meaning that the final anomaly score is calculated by averaging the anomaly scores of all the trees in the forest. More trees mean more accuracy but also more computational power required.

\begin{figure}[H]
	\includegraphics[width=6 cm]{img/IsolationForest.jpg}
	\caption{Isolation forest in action. The anomaly \textbf{x0} is isolated in less splits compared to the normal datapoint \textbf{x1}\cite{IsolationForestLiu}.}
	\label{fig:IsolationForest}
\end{figure}   
\unskip

The chosen implementation for isolation forest is the one provided by the \textit{Scikit-learn} library \cite{IsolationForestScikitLearn} for Python 3. 
The algorithm was tested on a Desktop PC with an Intel i7-6700 CPU and 16GB of RAM as well as on a Raspberry Pi 3 with a Quad Core 1.2GHz CPU and 1GB of RAM \cite{RaspberryPi3ModelB}.
The graphs shown are the ones obtained on the Desktop PC for convenience. The ability of the algorithm to run effectively on a resource constrained device like the Raspberry Pi 3, that has hardware comparable to the one found in high-end commercial drones, was confirmed.

\subsection{Evaluation Metrics}\label{EvaluationMetrics}

The performance of the model was evaluated using the state-of-the-art evaluation metrics for machine learning classification algorithms. The metrics are all based on the number of true positives, false positives, true negatives and false negatives. A true positive is an anomalous data point that is correctly classified as such, while a false positive is a normal data point that is incorrectly classified as an anomaly. A true negative on the other hand is a normal data point that is correctly classified as a normal data point, while a false negative is an anomalous data point which is incorrectly classified as a normal data point.
The evaluation metrics that we chose to use are the following:

\begin{itemize}
	\item \textbf{Accuracy:} The accuracy is the ratio of correctly classified data points to the total number of data points (equation \ref{eq:accuracy}).
	\begin{equation}
		\label{eq:accuracy}
	  Accuracy = \frac{TP + TN}{TP + TN + FP + FN}
	\end{equation}
	\item \textbf{Precision:} The precision is the ratio of correctly classified anomalies to the total number of data points classified as anomalies (equation \ref{eq:precision}).
	\begin{equation}
		\label{eq:precision}
	  Precision = \frac{TP}{TP + FP}
	\end{equation}
	\item \textbf{Recall:} The recall is the ratio of correctly classified anomalies to the total number of anomalies (equation \ref{eq:recall}).
	\begin{equation}
		\label{eq:recall}
	  Recall = \frac{TP}{TP + FN}
	\end{equation}
	\item \textbf{F1 Score:} The F1 score is the harmonic mean of the precision and recall (equation \ref{eq:f1}).
	\begin{equation}
		\label{eq:f1}
	  F1 = 2 \times \frac{Precision \times Recall}{Precision + Recall}
	\end{equation}
\end{itemize}

The combination of these metrics gives us an important insight into the model's performance.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

\subsection{Parameters Tuning Phase}

The first phase in our testing methodology was the tuning of the model's hyperparameters. The tuning was performed by making the parameters vary between a minimum and a maximum value, evaluating the impact on the performance metrics (depicted in section \ref{EvaluationMetrics}) and then choosing the best performing parameter value.
From the parameters available in the \textit{Scikit-learn} implementation of the Isolation Forest, the parameters we decided to tune were the following:

\begin{itemize}
	\item \textbf{n\_estimators:} The number of base estimators in the ensamble, i.e the number of isolation trees used to compute the anomaly score of each datapoint. 
	\item \textbf{max\_samples:} The max number of samples to draw from the dataset to train each tree.
	\item \textbf{contamination:} The amount of outliers present in the training dataset.  
\end{itemize}

During the tuning phase, the untested hyperparameters were set to their default values of the \textit{Scikit-learn} implementation of the Isolation Forest (\ref{tab:isolation_forest_parameters}).

\begin{table}[H]
	\caption{Scikit-Learn Isolation Forest hyperparameters default values.\label{tab:isolation_forest_parameters}}
	\newcolumntype{C}{>{\centering\arraybackslash}X}
	\begin{tabularx}{\textwidth}{CC}
	\toprule
	\textbf{Parameter} & \textbf{Default Value} \\
	\midrule
	n\_estimators & 100 \\
	max\_samples & 'auto' \\
	contamination & 0.1 \\
	\bottomrule
	\end{tabularx}
\end{table}

The model was evaluated using as input normal traffic samples concatenated with jamming attack samples (see code snippet \ref{alg:dataset_concatenation}). 
This was done to simulate a real-world scenario where the model has to be able to classify correctly anomalous samples but also reduce the number of false positives during normal operation.

\begin{algorithm}
	\caption{Test input definition}\label{alg:dataset_concatenation}
	\begin{algorithmic}[1]
	\State normalTraffic $\gets$ ReadAndParseFile(NORMAL\_TRAFFIC\_FILE, normal\_traffic\_size)
	\State jamming $\gets$ ReadAndParseFile(JAMMING\_FILE, jamming\_size)

	\State testInput $\gets$ Concatenate(normalTraffic, jamming)
	\end{algorithmic}
\end{algorithm}

The values whe chose for \textit{normal\_traffic\_size} and \textit{jamming\_size} are shown in table \ref{tab:dataset_sizes}.

\begin{table}[H]
	\caption{Dataset sizes used for the tuning and testing phases.}\label{tab:dataset_sizes}
	\newcolumntype{C}{>{\centering\arraybackslash}X}
	\begin{tabularx}{\textwidth}{CC}
	\toprule
	\textbf{Dataset} & \textbf{Size} \\
	\midrule
	normal\_traffic\_size & 20000 \\
	jamming\_size & 2000 \\
	\bottomrule
\end{tabularx}
\end{table}

The dataset is purposefully unbalanced between normal data points and anomalous points, as jamming attacks are usually rare events compared to normal network operation.

In figure \ref{fig:InputSignal} we can see a representation of the input signal in the case of a constant jamming attack. The proposed results, unless otherwise specified, are based on the input signal shown in figure \ref{fig:InputSignal}, which employs constant jamming. Periodic jamming always showed comparable trends as constant jamming in all the tests performed.

\begin{figure}[H]
	\begin{adjustwidth}{-\extralength}{0cm}
	\centering
	\includegraphics[width=19.5cm]{img/InputSignal.png}
	\caption{Input signal in the case of Constant Jamming. The normal traffic signal is concatenated with the jamming signal.}\label{fig:InputSignal}
\end{adjustwidth}
\end{figure}

\subsubsection{n\_estimators tuning}

The first parameter that we decided to tune was the \textit{n\_estimators} parameter. The tuning values are shown in table \ref{tab:n_estimators_tuning}.

\begin{table}[H]
	\caption{n\_estimators tuning values.}\label{tab:n_estimators_tuning}
	\newcolumntype{C}{>{\centering\arraybackslash}X}
	\begin{tabularx}{\textwidth}{CC}
	\toprule
	\textbf{n\_estimators} & \textbf{Values} \\
	\midrule
	Minimum & 1 \\
	Maximum & 50 \\
	Step & 1 \\
	\bottomrule
\end{tabularx}
\end{table}

The effect on the evaluation metrics of the model is shown in figure \ref{fig:n_estimators_tuning}.

\begin{figure}[H]
	\includegraphics[width=10.5cm]{img/nEstimatorsTuning.png}
	\caption{n\_estimators tuning results.}\label{fig:n_estimators_tuning}
\end{figure}
\unskip

As we can see from the graph, a growing number of estimators initially correlates with a better model performance, but after a certain point the performance stabilizes. 
In graphs \ref{fig:n_estimators_training_time} and \ref{fig:n_estimators_classification_time} is shown that the time required both for training and classification increases linearly with the number of estimators.
This means that choosing an appropriate number of estimators is crucial, as it can greatly affect the model performance.
